---
title: "Laptops"
author: "Augustus Chu"
date: '2025-02-26'
output: github_document
---

## Laptops Analysis

For this project, I will analyse an uncleaned dataset on Laptop Specs and their prices. The first section involves exploring, understanding then the cleaning the data for our purposes. The second section involves visualising the data. The main purpose of this project is to showcase what is possible with data analytics!

### Exploratory Analysis + Data Cleaning

```{r}
# Load packages
pacman::p_load(tidyverse, dplyr)
```

```{r}
# Read data
laptop.csv <- read_csv("laptopData.csv")
```

```{r}
# Analyse the data closely
(laptop.csv)
```

```{r}
# We can change column types to help us for the analysis/cleaning process
# We should also select only the important columns to keep
laptop <- laptop.csv %>%
  select(-`Unnamed: 0`) %>%
  mutate(Company = as_factor(Company),
         TypeName = as_factor(TypeName),
         Inches = as.double(Inches),
         Ram = as_factor(Ram),
         OpSys = as_factor(OpSys))

```

```{r}
# Lets take a look at the different brands of laptops we have. The bar graph shows us the most common brands are Lenovo, Dell, HP, Asus and Acer which all have over 100 laptops, with the remaining brands all having less than 100 laptops.
ggplot(laptop, aes(fct_infreq(Company))) + geom_bar()

# We also see there are some NA values, lets remove them.

laptop <- drop_na(laptop)

```

```{r}
# Now lets make some columns more usable, starting with the Screen Resolution info
# We can separate the actual resolution from the description, and change to a factor
laptop <- laptop %>%
  mutate(ScreenDescription = ScreenResolution,
    ScreenResolution = as_factor(str_split_i(ScreenResolution, " ", i=-1)))

# We can now see which size of screen is more common. There appears to be some niche screen sizes, but the majority of laptops have a screen size of 1920x1080 or 1366x768
ggplot(laptop, aes(fct_infreq(ScreenResolution))) + geom_bar()

```

```{r}
# Similarly, we can make the weight column more manageable
laptop <- laptop %>%
  mutate(Weight = round(as.numeric(str_sub(Weight, 1, -3)),2))

# We can now see what to expect for the weights of the laptops
ggplot(laptop, aes(Weight)) + geom_histogram()

```

```{r}
# The Price is currently in Indian Rupees, lets convert it to NZ Dollar, using a rate of 100Rupees : 2.04NZD
laptop <- laptop %>%
  mutate(PriceNZD = Price * 0.0204)

# Finally, lets select all the columns that have the most useful information
laptop <- laptop %>%
  select(Company, TypeName, Inches, ScreenResolution, Cpu, Ram, Memory, Gpu, OpSys, Weight, PriceNZD)

```

### Visualisations

Now that we have done our initial analysis, lets upgrade the visualisations of what we found

```{r}
# Firstly, we look at the different brands of laptops
ggplot(laptop, aes(fct_infreq(Company)),) + geom_bar(fill  = "#1D76B5", color = "white") + theme(legend.position = "bottom", axis.text.x = element_text(angle = 90)) + labs(x = "Company", y = "Count", title = "Laptop Brand")

```

```{r}
# Next, we look at the different weights of our laptops
ggplot(laptop, aes(Weight)) + geom_histogram(binwidth = 0.3, fill  = "#1D76B5", 
    color = "white") + labs(x = "Weight", y = "Count", title = "Weight of Laptops")

```

```{r}
# Perhaps we are interested in the prices of laptops for each company. Lets just focus on the 4 most common brands, and group the remaining ones. For this, we must group the laptop brands with low frequencies.

laptop <- laptop %>%
  mutate(CompanyGrouped = fct_lump(Company, n=4))

ggplot(laptop, aes(x = PriceNZD, y = CompanyGrouped, fill = CompanyGrouped)) + geom_boxplot() + scale_x_continuous(labels = scales::comma) + labs(x = "Price (NZD)", y = "", title = "Prices of Different Brands of Laptops") + guides(y = "none") +  scale_fill_discrete(name = "Company")

```

```{r}
# We can also find out which laptop is the cheapest or most expensive

laptop[which.min(laptop$PriceNZD),]

laptop[which.max(laptop$PriceNZD),]

```

```{r}
# Are larger laptops typically more expensive than smaller laptops? We can also compare different brands too
ggplot(laptop, aes(y = Weight, x = PriceNZD, col = CompanyGrouped)) + geom_point(alpha = 0.3, size = 3) + labs(title = "Price vs Weight of Laptops", x = "Price (NZD)") + guides(color = guide_legend(title = "Company")) 

```

Doing this type of analysis can help you find trends, or key points about the data which may bring valuable direction for making positive changes!
